---
title: "Data Manipulation"
output: github_document
---

```{r}
library(tidyverse)
```

## Load the FAS Litters data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```

## 'select' 

select lets us choose some columns and not others 

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
select(litters_df, -litter_number)
```

Renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmbeR = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LITTer_NUmbeR = litter_number)
```

Select helpers - help to quickly select

```{r}
select(litters_df, starts_with("gd"))
```

```{r}
select(litters_df, litter_number, everything())
```
these do the same thing ^
```{r}
relocate(litters_df, litter_number)
```

## 'filter' 

filter operates on rows while select operates on columns - you have to give filter a True vs False condition - use logical operators 

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

Logical ooperators for 'filter' 

```{r}
filter(litters_df, gd_of_birth == 20)
```

```{r}
filter(litters_df, !(gd_of_birth == 20))
```
! flips the logical operation

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Mod8")
```

```{r}
filter(litters_df, group %in% c("Mod8", "Con7"))
```

# 'mutate' 

allows us to modify an existing variable or create a new variable 

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```

```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```

## 'arrange' 

helps us put things "in order" 

```{r}
arrange(litters_df, pups_born_alive)
```

## %>% Pipe operation 

```{r}
litters_df =
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```


